---
title: "Project 3 - Predicitive Models"
author: "Smitali Paknaik and Paula Bailey"
date: "2022-11-04"
output: 
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    theme: readable
    df_print: "paged"
params:
  group:"lifestyle"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction section




  * The following packages are required for creating predictive models.

1. `dplyr` - A part of the `tidyverse` used for manipulating data
2. `tidyr` - A part of the `tidyverse` used for data cleaning
3. `ggplot2` - A part of the `tidyverse` used for creating graphics
4. 


```{r lib, include = FALSE}

library(tidyverse)
library(ggplot2)
library(purrr)
library(markdown)
library(caret)
library(GGally)     #create ggcorr and ggpairs plots

```

We used read.csv to load in the data.  The UCI site mentioned that `url` and `timedelta` are non-predictive variables, so we will remove them from our data set.  Afterwards, we parsed out the six lifestyle groups from the data set.

# Load Data, and check for NA
```{r}
data <- read.csv("OnlineNewsPopularity.csv") %>% select(-url, -timedelta)

#check for missing values
anyNA(data)  
```

For the automation, it will be easier if all the channels are in one column.  I used `pivot_longer()` to pivot columns: data_channel_is_lifestyle,data_channel_is_entertainment, data_channel_is_bus, data_channel_is_socmed, data_channel_is_tech, data_channel_is_world from wide to long format.
```{r}
dataPivot <- data %>% pivot_longer(cols = data_channel_is_lifestyle:data_channel_is_world, names_to = "channel",values_to = "Temp") 

head(dataPivot)
```
Now, the individual channel columns are combined into one column named channel.  The Temp value represents if an article exists for that particular change.  For the final data set, we will remove any values with 0.  The final data set has 33510 obs....**double check that you are not loosing too much data**  The original data set had 39644 obs, so we lost over 6000 observations... hmm.,, I'm fine with pivot, but not sure about removing missing information for day of the week.... again, come back.

```{r}
newData <- dataPivot %>% filter(Temp != 0) %>% select(-Temp)
```

These are my notes, so it seems like I'm jumping around.  I will perform pivot on "weekday" information as well. Afterwards, remove Temp2 which represents articles without a day of the week listed.  

```{r}
dataPivot1 <- newData %>% pivot_longer(cols = weekday_is_monday:weekday_is_sunday, names_to = "weekday",values_to = "Temp1") 
```


```{r}
newData <- dataPivot1 %>% filter(Temp1 != 0) %>% select(-Temp1)
```


```{r}
head(newData)
```


```{r}
colnames(newData)
```

```{r}
unique(newData$channel)
```


```{r}
# just for completing the analysis; I don't think we will need format once everything is automated correctly.  
lifestyle     <- newData %>% filter(channel == "data_channel_is_lifestyle")
entertainment <- newData%>% filter(channel == "data_channel_is_entertainment")
bus           <- newData%>% filter(channel == "data_channel_is_bus")
socmed        <- newData%>% filter(channel == "data_channel_is_socmed")
tech          <- newData%>% filter(channel == "data_channel_is_tech")
world         <- newData%>% filter(channel == "data_channel_is_world")
```


# Train and Test on practice - Lifestyle
```{r}

set.seed(21)
trainIndex <- createDataPartition(lifestyle$shares, p = 0.7, list = FALSE)
lifestyleTrain <- lifestyle[trainIndex, ]
lifestyleTest <- lifestyle[-trainIndex, ]

```


## Summaries on Training Data
```{r}
lifestyle %>% select(shares, starts_with("self"), starts_with("rate"), starts_with("avg") ) %>%summary(lifestyle)
```
Look at any others to see if their results should be included

We see the distribution of several quantitative variable  

    **Right-skewed** if its mean is **greater** than its median.    
    **Left-skewed** if its mean is **less** than its median.    
    **Normal** if its mean **equals** to its median.    


**Add contingency tables**
**Must have 3 graphs each of the data**

Share and positive words - hmmm, I expected something a bit more interesting....I figures there would be an uptick in articles shared.  Come back... JUst creating a bunch of plots to see what is interesting.
```{r}
ggplot(lifestyle, aes(x=rate_positive_words, y = shares ))+geom_point()
```

```{r}
ggplot(lifestyle, aes(x = shares ))+geom_boxplot(position = "dodge")
```

Share histogram
```{r}
ggplot(lifestyle, aes(x = shares ))+geom_histogram()
```
shares and global_sentiment_polarity
```{r}
ggplot(lifestyle, aes(x=num_videos, y = shares ))+geom_point()
```
We want to remove any predictor variables that are highly correlated to the response variable as called multicollinearity.  if variables have this characteristic it can lead to skewed or misleading results. We will create grouping of ten predictor variables to look at the relationship with our response variable share.

Looking at the results from the `ggcorr`, we do not see any highly correlated relationships.  if there was such relationship, it would be 1 or orange for highly positive correlated and -1 or blue for highly negative correlated.

**need to make corr plot look better - see if can change font on label**


```{r}
ggcorr(lifestyleTrain%>% dplyr::select(n_tokens_title, n_tokens_content, n_unique_tokens, n_non_stop_words, n_non_stop_unique_tokens, num_hrefs, num_self_hrefs, num_imgs, num_videos, shares),label_round = 2, label = "TRUE", label_size=3)
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select( average_token_length, num_keywords, shares),label_round = 2)
```


```{r}
ggcorr(lifestyleTrain%>% dplyr::select(kw_min_min, kw_max_min, kw_avg_min, kw_min_max, kw_max_max,  kw_avg_max, kw_min_avg, kw_max_avg, kw_avg_avg, shares),label_round = 2, label = "TRUE")
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select(self_reference_min_shares, self_reference_max_shares, self_reference_avg_sharess, shares),label_round = 2, label = "TRUE")
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select(LDA_00, LDA_01, LDA_02, LDA_03, LDA_04, global_subjectivity, global_sentiment_polarity, global_rate_positive_words, global_rate_negative_words, shares),label_round = 2, label = "TRUE")
```
```{r}
ggcorr(lifestyleTrain%>% dplyr::select(rate_positive_words, rate_negative_words, avg_positive_polarity, min_positive_polarity, max_positive_polarity,  avg_negative_polarity, min_negative_polarity, max_negative_polarity,  title_subjectivity, title_sentiment_polarity, abs_title_subjectivity, abs_title_sentiment_polarity, shares),label_round = 2, label = "TRUE")
```
# Varaible Selection
The final data set has 



# Linear Regerssion

Linear regression is a method to understand the relationship between a response variables Y and one or more predictor variables x, x1, x2, etc.  This method creates a line that best fits the data called "least known regression line"

For a Simple Linear Regression, we one response variable and one predictor variables. It uses the formula y = b0 + b1x, where....

For Multiple Linear Regression, we have one response variable and any number of predictor variable. It uses the formula Y = B0 + B1X1 + B2X2 + … + BpXp + E, where .....

Also add the 4 assumptions - linear relationship, independence, homoscedasticity, and Normality.


# Random Forest
Paula - Create RF





-----------------IGNORE for NOW----------------
Test area for params
to automatic Rmarkdown, see topic 2 notes, beginning on page 329: rewatch video

# Automation of data 
```{r}
#groups - need to consider a way to read in groups without hardcoding
socialID <- c("Lifestyle", "Entertainment","Business", "SocialMedia", "Tech", "World" )

#create filename
output_file <- paste0(socialID, "Analysis.html")
#create a group for each team with just the team name parameter
params = lapply(socialID, FUN = function(x){list(id = x)})

#put into a data frame
reports <- tibble(output_file, params)

library(rmarkdown)

apply(reports, MARGIN = 1,
      FUN = function(x){
        render(input = "./Project_3.Rmd",
               output_format = "github_document", 
               output_file = x[[1]], 
               params = x[[2]])
      })



```

```{r}
#::render("files/Project3.Rmd", output_file = "LifestyleAnalysis.html",
#params = list(team = "lifestyle"))
```

------------------------------------------------------------
## Smitali 
SOW

Introduction Section 
Summary statistics each member to show contribution
1- Linear Regression - each member 
Boosted Tree Model (with CV)
Any other support 