---
title: "Project 3 - Predicitive Models"
author: "Smitali Paknaik and Paula Bailey"
date: "2022-11-04"
output: 
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    theme: readable
    df_print: "paged"
params:
  group:"lifestyle"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction section




  * The following packages are required for creating predictive models.

1. `dplyr` - A part of the `tidyverse` used for manipulating data
2. `tidyr` - A part of the `tidyverse` used for data cleaning
3. `ggplot2` - A part of the `tidyverse` used for creating graphics
4. 


```{r lib, include = FALSE}
library(tidyverse)
library(ggplot2)
library(purr)
library(markdown)
library(caret)

```

We used read.csv to load in the data.  The UCI site mentioned that `url` and `timedelta` are non-predicitive variables, so we will remove them from our data set.  Afterwards, we parsed out the six lifestyle groups from the data set.

# Load Data and Parse Groups
```{r}
data <- read.csv("OnlineNewsPopularity.csv") %>% select(-url, -timedelta)

# just for completing the analysis; I don't think we will need this information once everything is automated correctly.
lifestyle     <- data[data$data_channel_is_lifestyle == 1,]
entertainment <- data[data$data_channel_is_entertainment == 1,]
bus           <- data[data$data_channel_is_bus == 1,]
socmed        <- data[data$data_channel_is_socmed == 1,]
tech          <- data[data$data_channel_is_tech == 1,]
world         <- data[data$data_channel_is_world == 1,]
```

# Train and Test on practice - Lifestyle
```{r}

set.seed(21)
trainIndex <- createDataPartition(lifestyle$share, p = 0.7, list = FALSE)
lifestyleTrain <- lifestyle[trainIndex, ]
lifestyleTest <- lifestyle[-trainIndex, ]

```

# Linear Regerssion
Paula - need to add short explanation on linear regression
Paula - summary on training data
Paula - also need to create individual LM


# Random Forest
Paula - Create RF





-----------------IGNORE for NOW----------------
Test area for params
to automatic Rmarkdown, see topic 2 notes, beginning on page 329: rewatch video
```{r}
#groups
socialID <- c("Lifestyle", "Entertainment","Business", "SocialMedia", "Tech", "World" )
#create filename
output_file <- paste0(socialID, "Analysis.html")
#create a group for each team with just the team name parameter
params = lapply(socialID, FUN = function(x){list(id = x)})

#put into a data frame
reports <- tibble(output_file, params)


```

```{r}
#::render("files/Project3.Rmd", output_file = "LifestyleAnalysis.html",
#params = list(team = "lifestyle"))
```

------------------------------------------------------------
