---
title: "Project 3 - Predicitive Models"
author: "Smitali Paknaik and Paula Bailey"
date: "2022-11-04"
output: 
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    theme: readable
    df_print: "paged"
params:
  group:"lifestyle"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction section




  * The following packages are required for creating predictive models.

1. `dplyr` - A part of the `tidyverse` used for manipulating data
2. `tidyr` - A part of the `tidyverse` used for data cleaning
3. `ggplot2` - A part of the `tidyverse` used for creating graphics
4. 


```{r lib, include = FALSE}

library(caret)      #training/test (splitting)
library(glmnet)     #best subset selection
library(GGally)     #create ggcorr and ggpairs plots
library(ggplot2)
library(leaps)      #identify different best models of different sizes
library(markdown)
library(MASS)       #access to forward and backward selection algorithms
library(purrr)
library(tidyverse)  #tidyverse set of packages and functions

```

We used read.csv to load in the data.  The UCI site mentioned that `url` and `timedelta` are non-predictive variables, so we will remove them from our data set.  Afterwards, I checked to validate the data set contained no missed values.  



# Load Data, and check for NA
```{r}
data <- read.csv("OnlineNewsPopularity.csv") %>% 
                              rename(`Lifestyle` =`data_channel_is_lifestyle`,
                                    Entertainment = `data_channel_is_entertainment`,
                                    Business      = `data_channel_is_bus`,
                                    SocialMedia = `data_channel_is_socmed`,
                                    Technology    = `data_channel_is_tech`,
                                    World       = `data_channel_is_world`,
                                    Monday      = `weekday_is_monday`,
                                    Tuesday     = `weekday_is_tuesday`,
                                    Wednesday   = `weekday_is_wednesday`,
                                    Thursday    = `weekday_is_thursday`,
                                    Friday      = `weekday_is_friday`,
                                    Saturday    = `weekday_is_saturday`,
                                    Sunday      = `weekday_is_sunday`) %>% 
                                    dplyr::select(-url, -timedelta)

#check for missing values
anyNA(data)  
```

For the automation, it will be easier if all the channels are in one column.  I used `pivot_longer()` to pivot columns: data_channel_is_lifestyle,data_channel_is_entertainment, data_channel_is_bus, data_channel_is_socmed, data_channel_is_tech, data_channel_is_world from wide to long format.
```{r}
dataPivot <- data %>% pivot_longer(cols = c("Lifestyle", "Entertainment", "Business", "SocialMedia", "Technology", "World"), names_to = "channel",values_to = "Temp") 

head(dataPivot)
```
Now, the individual channel columns are combined into one column named channel.  The Temp value represents if an article exists for that particular change.  For the final data set, we will remove any values with 0.  The final data set has 33510 obs....**double check that you are not loosing too much data**  The original data set had 39644 obs, so we lost over 6000 observations... hmm.,, I'm fine with pivot, but not sure about removing missing information for day of the week.... again, come back.

```{r}
newData <- dataPivot %>% filter(Temp != 0) %>% dplyr::select(-Temp)
```

These are my notes, so it seems like I'm jumping around.  I will perform pivot on "weekday" information as well. Afterwards, remove Temp2 which represents articles without a day of the week listed.  

```{r}
dataPivot1 <- newData %>% pivot_longer(c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), names_to = "weekday",values_to = "Temp1") 
```


```{r}
newData <- dataPivot1 %>% filter(Temp1 != 0) %>% dplyr::select(-Temp1)
```


```{r}
head(newData)
```


```{r}
colnames(newData)
```

```{r}
unique(newData$channel)
```


```{r}
# just for completing the analysis; I don't think we will need format once everything is automated correctly.  
lifestyle     <- newData %>% filter(channel == "Lifestyle")
entertainment <- newData%>% filter(channel == "Entertainment")
bus           <- newData%>% filter(channel == "Business")
socmed        <- newData%>% filter(channel == "SocailMedia")
tech          <- newData%>% filter(channel == "Technology")
world         <- newData%>% filter(channel == "World")
```


# Train and Test on practice - Lifestyle
```{r}

set.seed(21)
trainIndex <- createDataPartition(lifestyle$shares, p = 0.7, list = FALSE)
lifestyleTrain <- lifestyle[trainIndex, ]
lifestyleTest <- lifestyle[-trainIndex, ]

```


## Summaries on Training Data
```{r}
lifestyle %>% dplyr::select(shares, starts_with("rate"), starts_with("avg") ) %>%summary(lifestyle)
```
The distribution of several quantitative variables - shares, (response variable), rate_positive_words, rate_negative_words, avg_positive_polarity, and avg_negative_polarity.

We know the following about the distribution (shape):

    **Right-skewed** if mean is **greater** than median.    
    **Left-skewed** if mean is **less** than median.    
    **Normal** if mean **equals** to median.    

Share and positive words - hmmm, I expected something a bit more interesting....I figures there would be an uptick in articles shared.  Come back... JUst creating a bunch of plots to see what is interesting.
```{r}
ggplot(lifestyle, aes(x=rate_positive_words, y = shares ))+geom_point()
```

```{r}
ggplot(lifestyle, aes(x = shares ))+geom_boxplot(position = "dodge")
```

Share histogram
```{r}
ggplot(lifestyle, aes(x = shares ))+geom_histogram()
```
shares and global_sentiment_polarity
```{r}
ggplot(lifestyle, aes(x=num_videos, y = shares ))+geom_point()
```

## Contingency Tables

Articles per Channel vs Weekday
```{r}
table(newData$channel, newData$weekday)
```

Day of the Week vs Number of Keywords
```{r}
table(newData$weekday, newData$num_keywords)
```


## Correlations

We want to remove any predictor variables that are highly correlated to the response variable as called multicollinearity.  if variables have this characteristic it can lead to skewed or misleading results. We will create grouping of ten predictor variables to look at the relationship with our response variable share.

Looking at the results from the `ggcorr`, we do not see any highly correlated relationships.  if there was such relationship, it would be 1 or orange for highly positive correlated and -1 or blue for highly negative correlated.

**need to make corr plot look better - see if can change font on label**


```{r}
ggcorr(lifestyleTrain%>% dplyr::select(n_tokens_title, n_tokens_content, n_unique_tokens, n_non_stop_words, n_non_stop_unique_tokens, num_hrefs, num_self_hrefs, num_imgs, num_videos, shares),label_round = 2, label = "TRUE", label_size=3)
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select( average_token_length, num_keywords, shares),label_round = 2)
```


```{r}
ggcorr(lifestyleTrain%>% dplyr::select(kw_min_min, kw_max_min, kw_avg_min, kw_min_max, kw_max_max,  kw_avg_max, kw_min_avg, kw_max_avg, kw_avg_avg, shares),label_round = 2, label = "TRUE")
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select(self_reference_min_shares, self_reference_max_shares, self_reference_avg_sharess, shares),label_round = 2, label = "TRUE")
```

```{r}
ggcorr(lifestyleTrain%>% dplyr::select(LDA_00, LDA_01, LDA_02, LDA_03, LDA_04, global_subjectivity, global_sentiment_polarity, global_rate_positive_words, global_rate_negative_words, shares),label_round = 2, label = "TRUE")
```
```{r}
ggcorr(lifestyleTrain%>% dplyr::select(rate_positive_words, rate_negative_words, avg_positive_polarity, min_positive_polarity, max_positive_polarity,  avg_negative_polarity, min_negative_polarity, max_negative_polarity,  title_subjectivity, title_sentiment_polarity, abs_title_subjectivity, abs_title_sentiment_polarity, shares),label_round = 2, label = "TRUE")
```


# Varaible Selection

Before using any reduction algorithm to determine which variables to be in the model, I used the correlation tables above to reduce the predictors.  If the predictors are correlated, it's not necessary to similar predictor.

So I decided to remove min and max, if the predictor also contained age:
 min_positive_polarity
 max_positive_polarity
 min_negative_polarity
 max_negative_polarity
 self_reference_min_shares
 self_reference_max_shares
 kw_min_avg
 kw_max_avg
 kw_min_max
 kw_max_max
 kw_min_min
 kw_max_min
 
I removed all of the "LDA_".  When you view the correlation chart above, they seem to be correlated.  In addition, I am not sure what LDA means.  When I googled it, the results referenced modeling and Latent Dirichlet Allocation.
LDA_00: Closeness to LDA topic 0
LDA_01: Closeness to LDA topic 1
LDA_02: Closeness to LDA topic 2
LDA_03: Closeness to LDA topic 3
LDA_04: Closeness to LDA topic 4

I removed the following, because the data contains the absolute value of the same information:
title_subjectivity
title_sentiment_polarity

I removed the following, because of their strong relationship between global_rate_positive_words and global_rate_negative_words:
global_subjectivity
global_sentiment_polarity


I removed the following, because of their strong relationship between n_tokens_content:
n_unique_tokens
n_non_stop_unique_tokens
n_non_stop_words

The final data set will contain the following columns (features):
n_tokens_title
n_tokens_content
num_hrefs
num_self_hrefs
num_imgs
num_videos
average_token_length
num_keywords
kw_avg_min
kw_avg_max
kw_avg_avg
self_reference_avg_sharess
global_rate_positive_words
global_rate_negative_words
rate_positive_words
rate_negative_words
avg_positive_polarity
avg_negative_polarity
abs_title_subjectivity
abs_title_sentiment_polarity

So, that leaves us with 20. There is a relationship between num_hrefs and num_self_hrefs.  One of those may be removed later in the analysis.


```{r}
model1 <- lifestyle %>% dplyr::select(shares, n_tokens_title,
n_tokens_content,
num_hrefs,
num_self_hrefs,
num_imgs,
num_videos,
average_token_length,
num_keywords,
kw_avg_min,
kw_avg_max,
kw_avg_avg,
self_reference_avg_sharess,
global_rate_positive_words,
global_rate_negative_words,
rate_positive_words,
rate_negative_words,
avg_positive_polarity,
avg_negative_polarity,
abs_title_subjectivity,
abs_title_sentiment_polarity)
```

## Variable Selection

Even though we have reduced the predictors in the model using correlation, I decided to complete Best Selection on the final group of variables.  Come back - results do not look corret. review and delete later. Try using regsubset()


```{r}
trainControl = trainControl(method = "cv", number = 10)

bestFit <- train(shares ~ ., data = model1,
method = 'leapSeq',
preProcess = c("center", "scale"),
tuneGrid  = data.frame(nvmax = 1:20),
trControl = trainControl
)

bestFit$results
```

```{r}
bestFit$bestTune

```

```{r}
summary(bestFit$finalModel)
```

```{r}
coef(bestFit$finalModel, 6)
```

# Linear Regression

Linear regression is a method to understand the relationship between a response variables Y and one or more predictor variables x, x1, x2, etc.  This method creates a line that best fits the data called "least known regression line"

For a Simple Linear Regression, we one response variable and one predictor variables. It uses the formula y = b0 + b1x, where....

For Multiple Linear Regression, we have one response variable and any number of predictor variable. It uses the formula Y = B0 + B1X1 + B2X2 + … + BpXp + E, where .....

Also add the 4 assumptions - linear relationship, independence, homoscedasticity, and Normality.


I will split the data again, since I have a final data set of 21

```{r}
set.seed(21)
trainIndexSub <- createDataPartition(model1$shares, p = 0.7, list = FALSE)
model1Train <- lifestyle[trainIndex, ]
model1Test <- lifestyle[-trainIndex, ]
```

**Trying to figure out which variable is not a factor (each observation has the same value)**
Error in `contrasts<-`(`*tmp*`, value = contr.funs[1 + isOF[nn]]) : 
contrasts can be applied only to factors with 2 or more levels

```{r}
sapply(lapply(model1, unique), length)
```
```{r}
#display unique values for each variable
lapply(model1[c('num_self_hrefs', 'num_keywords', 'num_imgs')], unique)
```


```{r}
regr1 <- regsubsets(shares ~., data = model1Train, nvmax=20)
summary(regr1)

```


# Random Forest
Paula - Create RF





-----------------IGNORE for NOW----------------
Test area for params
to automatic Rmarkdown, see topic 2 notes, beginning on page 329: rewatch video

# Automation of data 
```{r}
#groups - need to consider a way to read in groups without hardcoding
socialID <- c("Lifestyle", "Entertainment","Business", "SocialMedia", "Tech", "World" )

#create filename
output_file <- paste0(socialID, "Analysis.html")
#create a group for each team with just the team name parameter
params = lapply(socialID, FUN = function(x){list(id = x)})

#put into a data frame
reports <- tibble(output_file, params)

library(rmarkdown)

apply(reports, MARGIN = 1,
      FUN = function(x){
        render(input = "./Project_3.Rmd",
               output_format = "github_document", 
               output_file = x[[1]], 
               params = x[[2]])
      })



```

```{r}
#::render("files/Project3.Rmd", output_file = "LifestyleAnalysis.html",
#params = list(team = "lifestyle"))
```

------------------------------------------------------------
## Smitali 
SOW

Introduction Section 
Summary statistics each member to show contribution
1- Linear Regression - each member 
Boosted Tree Model (with CV)
Any other support 